### 2015-02-15 Data/kernels.Rmd

---
title: "Density estimations for dispersal kernels"
author: "Pedro Jordano"
date: "enero 29, 2015"
output: html_document
bibliography: refs.bib
---

Code for ploting diffreent forms of dispersal kernels and testing dispersal 
model functions.

Data refer to seed dispersal distances estimated by direct assignement from
seed endocarp genotypes

#### Plotting the kernels

```{r plain, fig.height=5, fig.width=8}
# Plain frequency distribution of dispersal distances.
require(MASS)
x<-assdist$dist
truehist(x,xlim=c(0,20000),prob=F,h=100,col=1,xlab="Distance (m)",ylab="Probability",ymax=350)
rug(x,side=1,col="red")

```

#### Fitting models

```{r, echo=FALSE}

```

----------------------------------
```{r}
sessionInfo()
```
----------------------------------

### References